\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-14 chapters/stbf_struct.tex
Started processing comments MDV

  There are multiple state-of-the-art P3 classification methods, like
  \acp{svm}~\cite{Tayeb2014}, deep
	learning models~\cite{Vareka2020,Borra2020}, and Riemannian Geometry
	classifiers~\cite{Barachant2014}.
	While these models often return a high classification accuracy, there is a need
	for lightweight models -- lightweight models lead to fast off-line analyses and can be transferred to consumer-grade hardware.
	When moving towards plug-and-play solutions, \ac{bci} calibration sessions should be short and model training times low.
  The \ac{stbf}~\cite{VanVliet2015, Wittevrongel2016} belongs to
	this class of \ac{erp} decoding models as it achieves state-of-the-art performance and is fast to train.
	Earlier work has shown that it is
	possible to apply the spatiotemporal beamformer to multiple time-locked visual
  \ac{bci} paradigms, including the P3 oddball paradigm, \ac{ssvep},
  \ac{cvep}~\cite{Wittevrongel2017a}, \ac{mvep}~\cite{Libert2021}.

	This work shows that the original spatiotemporal
	beamformer~\cite{Wittevrongel2016} can fall short in performance when \ac{bci}
	calibration data are restricted.
	We also show that the spatiotemporal beamformer does not scale well for
	higher spatial and temporal resolution cases.
	As a response to these issues, we introduce a regularization method that
	exploits prior knowledge about the spatiotemporal nature of the \ac{eeg} signal to
	improve the accuracy for low data availability settings and speed up the
	classifier training time, thereby considerably reducing memory usage.
	Similar structured regularization approaches have been applied to other linear
	\ac{erp} classifiers~\cite{GonzalezNavarro2017, Vliet2020} and have shown
	significant increases in performance.
	Additionally, we show that regularization results in an interpretable
	classification model, which can aid in analyzing and developing spatiotemporal beamformer-based classifiers.

@@

	\subsection{Spatiotemporal beamforming}
  \ac{lcmv}-beamforming was initially introduced to \ac{eeg} analysis as a filter for
	source localization~\cite{VanVeen1997} to enhance the signal-to-noise ratio
	(SNR).
	\textcite{VanVliet2015} first applied the spatiotemporal
	\ac{lcmv}-beamformer as a method for the analysis of \acp{erp}.
	The extension to the combined spatiotemporal domain~\cite{VanVliet2015} and the
	data-driven approaches proposed by Treder et al.~\cite{Treder2016} and
	Wittevrongel et al.~\cite{Wittevrongel2016} allow for its application to classification problems.

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:minimum_variance} under the linear constraint given by
	\cref{eq:linear_constraint} returns the filter weights $w$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \underset{\mat{w}}{\arg\min}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:minimum_variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:linear_constraint}
	\end{equation}
	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \text{vec}(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
	epochs, determined as follows:
	\begin{equation}
		\mathbf{a} =
		\frac{1}{|\text{class 1}|}\sum_\text{class 1}\mathbf{x_i} -
		\frac{1}{|\text{class 2}|}\sum_\text{class 2}\mathbf{x_i}
		\label{eq:activation_pattern}
	\end{equation}

	The method of Lagrange multipliers then gives the closed-form solution to the minimization problem posed by
	\cref{eq:minimum_variance} and \cref{eq:linear_constraint} as:
	\begin{equation}
		\mat{w} =
    \frac{\mat{C}^{-1}\mat{a}^\intercal}
		{\mat{a}C^{-1}\mat{a}^\intercal}
		\label{eq:closed_form}
	\end{equation}
	The beamformer can be applied to epochs (unseen or not) as:
	\begin{equation}
		y_i = \mathbf{w}\mathbf{x}_i
		\label{eq:apply_beamformer}
	\end{equation}
	resulting in a scalar output per epoch.
	The linear constraint in \cref{eq:linear_constraint} ensures that the
	beamformer maps epochs containing a target response to a score close to one
	and, conversely, epochs not containing a target response to a score close to
	zero.

@@

	\subsubsection{Empirical covariance estimation}
	\label{sec:empirical_covariance}
	Earlier spatiotemporal beamformer studies~\cite{Wittevrongel2016,
		Wittevrongel2016a, Wittevrongel2017, Wittevrongel2017a} use the empirical
	covariance and inverse covariance calculated as follows:
	\begin{equation}
    \hat{\mat{C}}_\text{emp} =
		\frac{1}{N-1}\sum^{N}_{i=1}\mat{x}_i\mat{x}_i^\intercal
		\label{eq:empirical_covariance}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{emp} = \hat{\mat{C}}_\text{emp}^+
		\label{eq:empirical_inverse_covariance}
	\end{equation}
	The Moore-Penrose pseudoinverse $^+$, ensures a solution exists when
  $\hat{\mat{C}}_\text{emp}$ is singular.
	\cref{fig:kronlda-covs}a and \cref{fig:kronlda-covs}b respectively show examples of the
	empirical estimators of the covariance and the inverse covariance matrices.
	The empirical estimator suffers from performance and
	stability issues if the number of epochs $N$ used or estimation is not much larger than the number of features $CS$~\cite{Stein1956,Khatri1987}.

	\subsubsection{Shrunk covariance estimation}
	\label{sec:shrunk_covariance}
	The shrinkage covariance estimator creates a better-conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:
	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\text{Tr}(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:shrinkage_inverse}
	\end{equation}
	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

@@

	We opt for the \ac{loocv} shrinkage estimator because it avoids some of the
	assumptions made by~\cite{Ledoit2004} and~\cite{Chen2010} and
	because it generalizes to structured covariance estimation as described in
	\cref{seq:structured_estimation}.

	\subsubsection{Spatiotemporal beamforming with Kronecker-Toeplitz structured covariance}
	\label{seq:structured_estimation}
	Exploiting prior knowledge about the spatiotemporal structure of the \ac{eeg} signal leads to a more regularized estimator of the covariance.
	When viewing the example of empirical spatiotemporal \ac{eeg} covariance in
	\cref{fig:kronlda-covs}a, it becomes clear that this matrix consists of a block pattern of repeated, similar matrices.
	Due to the multi-channel nature of the signal, we assume that the covariance of spatiotemporal \ac{eeg} epochs is a Kronecker
	product of two smaller
	matrices~\cite{Munck1992,DeMunck1999,Huizenga2002}, as expressed
	by:
	\begin{equation}
    \hat{\mat{C}}_\text{struct} = \hat{\mat{S}} \otimes \hat{\mat{T}}
		\label{eq:kronecker}
	\end{equation}
	with $\otimes$ the Kronecker product operator.
  $\hat{\mat{S}} \in \mathbb{R}^{C\times C}$ and $\hat{\mat{T}} \in \mathbb{R}^{S\times S}$ respectively correspond to estimators of the spatial and temporal covariance of the data.
	Furthermore, because the temporal covariance of the \ac{eeg}-signal is
	stationary (i.e., it is only dependent on interval length between covarying
	time samples)~\cite{Bijma2003}, it is assumed to have a Toeplitz-matrix structure:
	\begin{equation}
    \hat{\mat{T}}_{i,j} = \hat{\mat{T}}_{i+1,j+1}
		\label{eq:toeplitz}
	\end{equation}
	\Cref{prop:inverse_kronecker} then leads to
	\cref{eq:cov_inverse_kronecker} to estimate the inverse
	covariance.

	\begin{property}
    $(\mat{U} \otimes \mat{V})^+ = \mat{U}^+ \otimes \mat{V}^+$ for any non-singular
    matrices $\mat{U}$ and $\mat{V}$~\cite{Langville2004}.
		\label{prop:inverse_kronecker}
	\end{property}

	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{struct} = \hat{\mat{S}}^+ \otimes \hat{\mat{T}}^+
		\label{eq:cov_inverse_kronecker}
	\end{equation}

	Finally, based on \cref{prop:kronecker_multiplication},
	\cref{eq:closed_form} can be reformulated more efficiently as
	\cref{eq:closed_form_kronecker}.
	\begin{property}
    $(\mat{U}\otimes \mat{V})\cdot\text{vec}(\mat{W}) =
    \text{vec}(\mat{V}\mat{W}\mat{U}^\intercal)$
    for any matrices $\mat{U}\in\mathbb{R}^{P\times P}$,
    $\mat{V}\in\mathbb{R}^{Q\times Q}$ and $\mat{W}\in\mathbb{R}^{P\times Q}$~\cite{Loan2000}.
		\label{prop:kronecker_multiplication}
	\end{property}

	\begin{equation}
		\hat{\mat{w}}_\text{struct} =
    \frac{\hat{\mat{S}}^+\mat{A}^\intercal\hat{\mat{T}}^+}
    {\mat{a}\cdot\text{vec}(\hat{\mat{S}}^+A^\intercal\hat{\mat{T}}^+)}
		\label{eq:closed_form_kronecker}
	\end{equation}

	Using \cref{eq:closed_form_kronecker} removes the need to calculate the
  full, high dimensional Kronecker product $\hat{\mat{S}}^+\otimes
    \hat{\mat{T}}^+$.
	\Cref{fig:kronlda-covs}e and \cref{fig:kronlda-covs}f respectively show examples of the
	structured covariance and inverse covariance estimators,
	consisting of a spatial Kronecker factor (\cref{fig:kronlda-covs}g and
	\cref{fig:kronlda-covs}h) and a temporal component (\cref{fig:kronlda-covs}i and
	\cref{fig:kronlda-covs}j).

@@

	\subsubsection{Kronecker-Toeplitz structured covariance estimation}
	\label{sec:structured_covariance}
	The question remains how to estimate $\hat{S}$ and $\hat{T}$.
	While the Flip-Flop and Non-iterative Flip-Flop
	algorithms~\cite{Lu2005, Werner2008, Wirfaelt2010} can estimate Kronecker or Kronecker-Toeplitz structured covariances, new results show that a fixed point iteration is more efficient~\cite{Wiesel2012a,Wiesel2012}.
	After each iteration, the spatial and temporal covariances matrices are scaled to unit
	variance to ensure the fixed point iteration converges.
	Finally, shrinkage can also be introduced in the Fixed Point Iteration to
	improve stability and achieve more robust
	regularization~\cite{Wiesel2012,Greenewald2014,Beltrachini2013, Breloy2016}.
	The spatial and temporal covariance matrices are shrunk at every fixed-point
	iteration with shrinkage factors $\beta_k$ and $\gamma_k$ before matrix
	inversion in the
	next iteration.
	Combined, this leads to the iterative estimation algorithm described by the
	following equations:

	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1} =
			\frac{1}{N}
      \sum^N_{i=1}\mat{X}_i^\intercal\hat{\mat{T}}_k^+\mat{X}_i
			\label{eq:fpi_spatial}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1} =
			\frac{1}{N}
      \sum^N_{i=1}\mat{X}_i\hat{\mat{S}}_k^+\mat{X}_i^\intercal
			\label{eq:fpi_temporal}
		\end{equation}
	\end{subequations}

	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1}^{(\beta)} =
      (1-\beta_{k+1})\tilde{\mat{S}}_{k+1}
      +\beta_{k+1}\frac{\text{Tr}(\tilde{\mat{S}}_{k+1})}{C}\mathbb{I}
			\label{eq:fpi_spatial_shrunk}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1}^{(\gamma)} =
      (1-\gamma_{k+1})\tilde{\mat{T}}_{k+1}
      +\gamma_{k+1}\frac{\text{Tr}(\tilde{\mat{T}}_{k+1})}{S}\mathbb{I}
			\label{eq:fpi_temporal_shrunk}
		\end{equation}
	\end{subequations}

	\begin{subequations}
		\begin{equation}
      \hat{\mat{S}}_{k+1} =
      \frac{C}{\text{Tr}\left[\tilde{\mat{S}}_{k+1}^{(\beta)}\right]}
      \tilde{\mat{S}}_{k+1}^{(\beta)}
			\label{eq:fpi_spatial_norm}
		\end{equation}
		\begin{equation}
      \hat{\mat{T}}_{k+1} =
      \frac{S}{\text{Tr}\left[\tilde{\mat{T}}_{k+1}^{(\gamma)}\right]}
      \tilde{\mat{T}}_{k+1}^{(\gamma)}
			\label{eq:fpi_temporal_norm}
		\end{equation}
	\end{subequations}

  $\hat{\mat{S}}_0$ and $\hat{\mat{T}}_0$ can be initialized to any positive definite matrix.
	We choose to use the identity matrices $\mathbb{I}^{C\times C}$ and $\mathbb{I}^{S\times S}$.
  After each iteration, all diagonals of $\hat{\mat{R}}_{k+1}$ are set to their mean
  values to ensure that $\hat{\mat{R}}_{k+1}$ and $\hat{\mat{T}}_{k+1}$ are Toeplitz structured.

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\hat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}
			\label{eq:spatial_shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
				+ \frac{1}{S}\text{Tr}(\tilde{T}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\mat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\hat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}
			\label{eq:temporal_shrinkage}
		\end{equation}
	\end{subequations}
	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

	%	identity matrix and assuming identity covariance in \cref{eq:closed_form} yields weights

>>>>

  There are multiple state-of-the-art P3 classification methods, like
  \acp{svm}~\cite{Tayeb2014}, deep
	learning models~\cite{Vareka2020,Borra2020}, and Riemannian Geometry
	classifiers~\cite{Barachant2014}.
	While these models often return a high classification accuracy, there is a need
	for lightweight models -- lightweight models lead to fast off-line analyses and can be transferred to consumer-grade hardware.
	When moving towards plug-and-play solutions, \ac{bci} calibration sessions should be short and model training times low.
  The \ac{stbf}~\cite{VanVliet2015, Wittevrongel2016} belongs to
	this class of \ac{erp} decoding models as it achieves state-of-the-art performance and is fast to train.
	Earlier work has shown that it is
	possible to apply the spatiotemporal beamformer to multiple time-locked visual
  \ac{bci} paradigms, including the P3 oddball paradigm, \ac{ssvep},
  \ac{cvep}~\cite{Wittevrongel2017a}, and the \ac{mvep}~\cite{Libert2021}.

	This work shows that the original spatiotemporal
	beamformer~\cite{Wittevrongel2016} can fall short in performance when \ac{bci}
	calibration data are restricted.
	We also show that the spatiotemporal beamformer does not scale well for
	higher spatial and temporal resolution cases.
	As a response to these issues, we introduce a regularization method that
	exploits prior knowledge about the spatiotemporal nature of the \ac{eeg} signal to
	improve the accuracy for low data availability settings and speed up the
	classifier training time, thereby considerably reducing memory usage.
	Similarly structured regularization approaches have been applied to other linear
	\ac{erp} classifiers~\cite{GonzalezNavarro2017, Vliet2020} and have shown
	significant increases in performance.
%	Additionally, we show that regularization results in an interpretable
%	classification model, which can aid in analyzing and developing spatiotemporal beamformer-based classifiers.

@@

	\subsection{Spatiotemporal beamforming}
  \ac{lcmv}-beamforming was initially introduced to \ac{eeg} analysis as a filter for
	source localization~\cite{VanVeen1997} to enhance the signal-to-noise ratio
	(SNR).
	\textcite{VanVliet2015} first applied the spatiotemporal
	\ac{lcmv}-beamformer as a method for the analysis of \acp{erp}.
	The extension to the combined spatiotemporal domain~\cite{VanVliet2015} and the
	data-driven approaches proposed by \textcite{Treder2016} and
	Wittevrongel et al.~\cite{Wittevrongel2016} allow for its application to classification problems.

	For the following analyses, we assume that all \ac{eeg} channels are normalized with zero mean and unit variance without loss of generality.
	Solving \cref{eq:stbf-struct/minimum-variance} under the linear constraint given by
  \cref{eq:stbf-struct/linear-constraint} returns the filter weights $\mat{w}$ defining the spatiotemporal \ac{lcmv}-beamformer.
	\begin{equation}
    \argmin_\mat{w}\mat{w}^\intercal \mat{C}
		\mat{w}^\intercal
		\label{eq:stbf-struct/minimum-variance}
	\end{equation}
	\begin{equation}
		\mat{a}^\intercal\mat{w} = 1
		\label{eq:stbf-struct/linear-constraint}
	\end{equation}
	These weights minimize the variance of the output of the filter while enhancing
	the signal characterized by the constraint.
  $\mat{a} = \text{vec}(\mat{A})$ is the data-driven activation pattern, a template
	of the signal of interest maximizing the difference between two classes of
	epochs, determined as follows:
	\begin{equation}
		\mathbf{a} =
		\frac{1}{|\text{class 1}|}\sum_\text{class 1}\mathbf{x_i} -
		\frac{1}{|\text{class 2}|}\sum_\text{class 2}\mathbf{x_i}
		\label{eq:stbf-struct/ap}
	\end{equation}

	The method of Lagrange multipliers then gives the closed-form solution to the minimization problem posed by
	\cref{eq:stbf-struct/minimum-variance} and \cref{eq:stbf-struct/linear-constraint} as:
	\begin{equation}
		\mat{w} =
    \frac{\mat{C}^{-1}\mat{a}^\intercal}
    {\mat{a}\mat{C}^{-1}\mat{a}^\intercal}
		\label{eq:stbf-struct/closed-form}
	\end{equation}
	The beamformer can be applied to epochs (unseen or not) as:
	\begin{equation}
		y_i = \mathbf{w}\mathbf{x}_i
		\label{eq:stbf-struct/apply-bf}
	\end{equation}
	resulting in a scalar output per epoch.
	The linear constraint in \cref{eq:stbf-struct/linear-constraint} ensures that the
	beamformer maps epochs containing a target response to a score close to one
	and, conversely, epochs not containing a target response to a score close to
	zero.

@@

	\subsubsection{Empirical covariance estimation}
	\label{sec:stbf-struct/methods/emp-cov}
	Earlier spatiotemporal beamformer studies~\cite{Wittevrongel2016,
		Wittevrongel2016a, Wittevrongel2017, Wittevrongel2017a} use the empirical
	covariance and inverse covariance calculated as follows:
	\begin{equation}
    \hat{\mat{C}}_\text{emp} =
		\frac{1}{N-1}\sum^{N}_{i=1}\mat{x}_i\mat{x}_i^\intercal
		\label{eq:stbf-struct/emp-cov}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{emp} = \hat{\mat{C}}_\text{emp}^+
		\label{eq:stbf-struct/emp-inv-cov}
	\end{equation}
	The Moore-Penrose pseudoinverse $^+$, ensures a solution exists when
  $\hat{\mat{C}}_\text{emp}$ is singular.
	\cref{fig:kronlda-covs}a and \cref{fig:kronlda-covs}b respectively show examples of the
	empirical estimators of the covariance and the inverse covariance matrices.
	The empirical estimator suffers from performance and
	stability issues if the number of epochs $N$ used or estimation is not much larger than the number of features $CS$~\cite{Stein1956,Khatri1987}.

	\subsubsection{Shrunk covariance estimation}
	\label{sec:stbf-struct/methods/shrunk-cov}
	The shrinkage covariance estimator creates a better-conditioned inversion matrix problem and generally performs better when applied to unseen data.
	The estimators for the covariance and inverse covariance are given by:
	\begin{equation}
    \hat{\mat{C}}_\alpha =
    (1-\alpha) \hat{\mat{C}}_\text{emp}
    + \alpha\frac{\text{Tr}(\hat{\mat{C}}_\text{emp})}{CS}\mathbb{I}
		\label{eq:shrinkage}
	\end{equation}
	\begin{equation}
    \widehat{\mat{C}^{-1}}_\alpha =
    \hat{\mat{C}}^+_\alpha
		\label{eq:stbf-struct/shrinkage-inv}
	\end{equation}
	with $0<\alpha<1$.
	Analogous to $L_2$ regularization of the beamforming problem,
	shrinkage reduces the ratio between the smallest and largest eigenvalues
	of the covariance matrix by strengthening the diagonal.
	\cref{fig:kronlda-covs}c and \cref{fig:kronlda-covs}d respectively show examples of the
	shrunk estimator of the covariance and the inverse covariance matrices.

@@

	We opt for the \ac{loocv} shrinkage estimator because it avoids some of the
	assumptions made by~\cite{Ledoit2004} and~\cite{Chen2010} and
	because it generalizes to structured covariance estimation as described in
	\cref{sec:stbf-struct/methods/structured-estimation}.

	\subsubsection{Spatiotemporal beamforming with Kronecker-Toeplitz structured covariance}
	\label{sec:stbf-struct/methods/structured-estimation}
	Exploiting prior knowledge about the spatiotemporal structure of the \ac{eeg} signal leads to a more regularized estimator of the covariance.
	When viewing the example of empirical spatiotemporal \ac{eeg} covariance in
	\cref{fig:kronlda-covs}a, it becomes clear that this matrix consists of a block pattern of repeated, similar matrices.
	Due to the multi-channel nature of the signal, we assume that the covariance of spatiotemporal \ac{eeg} epochs is a Kronecker
	product of two smaller
	matrices~\cite{Munck1992,DeMunck1999,Huizenga2002}, as expressed
	by:
	\begin{equation}
    \hat{\mat{C}}_\text{struct} = \hat{\mat{S}} \otimes \hat{\mat{T}}
		\label{eq:kronecker}
	\end{equation}
	with $\otimes$ the Kronecker product operator.
  $\hat{\mat{S}} \in \mathbb{R}^{C\times C}$ and $\hat{\mat{T}} \in \mathbb{R}^{S\times S}$ respectively correspond to estimators of the spatial and temporal covariance of the data.
	Furthermore, because the temporal covariance of the \ac{eeg}-signal is
	stationary (i.e., it is only dependent on interval length between covarying
	time samples)~\cite{Bijma2003}, it is assumed to have a Toeplitz-matrix structure:
	\begin{equation}
    \hat{\mat{T}}_{i,j} = \hat{\mat{T}}_{i+1,j+1}
		\label{eq:toeplitz}
	\end{equation}
	\Cref{prop:stbf-struct/inverse-kronecker} then leads to
	\cref{eq:stbf-struct/cov-inverse-kronecker} to estimate the inverse
	covariance.

	\begin{property}
    $(\mat{U} \otimes \mat{V})^+ = \mat{U}^+ \otimes \mat{V}^+$ for any non-singular
    matrices $\mat{U}$ and $\mat{V}$~\cite{Langville2004}.
		\label{prop:stbf-struct/inverse-kronecker}
	\end{property}

	\begin{equation}
    \widehat{\mat{C}^{-1}}_\text{struct} = \hat{\mat{S}}^+ \otimes \hat{\mat{T}}^+
		\label{eq:stbf-struct/cov-inverse-kronecker}
	\end{equation}

	Finally, based on \cref{propr:stbf-struct/kron-multiplication},
	\cref{eq:stbf-struct/closed-form} can be reformulated more efficiently as
	\cref{eq:stbf-struct/closed-form-kron}.
	\begin{property}
    $(\mat{U}\otimes \mat{V})\cdot\text{vec}(\mat{W}) =
    \text{vec}(\mat{V}\mat{W}\mat{U}^\intercal)$
    for any matrices $\mat{U}\in\mathbb{R}^{P\times P}$,
    $\mat{V}\in\mathbb{R}^{Q\times Q}$ and $\mat{W}\in\mathbb{R}^{P\times Q}$~\cite{Loan2000}.
		\label{propr:stbf-struct/kron-multiplication}
	\end{property}

	\begin{equation}
		\hat{\mat{w}}_\text{struct} =
    \frac{\hat{\mat{S}}^+\mat{A}^\intercal\hat{\mat{T}}^+}
    {\mat{a}\cdot\text{vec}(\hat{\mat{S}}^+A^\intercal\hat{\mat{T}}^+)}
		\label{eq:stbf-struct/closed-form-kron}
	\end{equation}

	Using \cref{eq:stbf-struct/closed-form-kron} removes the need to calculate the
  full, high dimensional Kronecker product $\hat{\mat{S}}^+\otimes
    \hat{\mat{T}}^+$.
	\Cref{fig:kronlda-covs}e and \cref{fig:kronlda-covs}f respectively show examples of the
	structured covariance and inverse covariance estimators,
	consisting of a spatial Kronecker factor (\cref{fig:kronlda-covs}g and
	\cref{fig:kronlda-covs}h) and a temporal component (\cref{fig:kronlda-covs}i and
	\cref{fig:kronlda-covs}j).

@@

	\subsubsection{Kronecker-Toeplitz structured covariance estimation}
	\label{sec:stbf-struct/methods/struct-cov}
  The question remains how to estimate $\mat{\hat{S}}$ and $\mat{\hat{T}}$.
	While the Flip-Flop and Non-iterative Flip-Flop
	algorithms~\cite{Lu2005, Werner2008, Wirfaelt2010} can estimate Kronecker or Kronecker-Toeplitz structured covariances, new results show that a fixed point iteration is more efficient~\cite{Wiesel2012a,Wiesel2012}.
	After each iteration, the spatial and temporal covariances matrices are scaled to unit
	variance to ensure the fixed point iteration converges.
	Finally, shrinkage can also be introduced in the Fixed Point Iteration to
	improve stability and achieve more robust
	regularization~\cite{Wiesel2012,Greenewald2014,Beltrachini2013, Breloy2016}.
	The spatial and temporal covariance matrices are shrunk at every fixed-point
	iteration with shrinkage factors $\beta_k$ and $\gamma_k$ before matrix
	inversion in the
	next iteration.

	Combined, this leads to the iterative estimation algorithm described by the
	following equations:
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1} =
			\frac{1}{N}
      \sum^N_{i=1}\mat{X}_i^\intercal\hat{\mat{T}}_k^+\mat{X}_i
			\label{eq:stbf-struct/fpi-spatial}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1} =
			\frac{1}{N}
      \sum^N_{i=1}\mat{X}_i\hat{\mat{S}}_k^+\mat{X}_i^\intercal
			\label{eq:stbf-struct/fpi-temporal}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \tilde{\mat{S}}_{k+1}^{(\beta)} =
      (1-\beta_{k+1})\tilde{\mat{S}}_{k+1}
      +\beta_{k+1}\frac{\text{Tr}(\tilde{\mat{S}}_{k+1})}{C}\mathbb{I}
			\label{eq:stbf-struct/fpi-spatial-shrunk}
		\end{equation}
		\begin{equation}
      \tilde{\mat{T}}_{k+1}^{(\gamma)} =
      (1-\gamma_{k+1})\tilde{\mat{T}}_{k+1}
      +\gamma_{k+1}\frac{\text{Tr}(\tilde{\mat{T}}_{k+1})}{S}\mathbb{I}
			\label{eq:stbf-struct/fpi-temporal-shrunk}
		\end{equation}
	\end{subequations}
	\begin{subequations}
		\begin{equation}
      \hat{\mat{S}}_{k+1} =
      \frac{C}{\text{Tr}\left[\tilde{\mat{S}}_{k+1}^{(\beta)}\right]}
      \tilde{\mat{S}}_{k+1}^{(\beta)}
			\label{eq:stbf-struct/fpi-spatial-norm}
		\end{equation}
		\begin{equation}
      \hat{\mat{T}}_{k+1} =
      \frac{S}{\text{Tr}\left[\tilde{\mat{T}}_{k+1}^{(\gamma)}\right]}
      \tilde{\mat{T}}_{k+1}^{(\gamma)}
			\label{eq:stbf-struct/fpi-temporal-norm}
		\end{equation}
	\end{subequations}
  $\hat{\mat{S}}_0$ and $\hat{\mat{T}}_0$ can be initialized to any positive definite matrix.
	We choose to use the identity matrices $\mathbb{I}^{C\times C}$ and $\mathbb{I}^{S\times S}$.
  After each iteration, all diagonals of $\hat{\mat{R}}_{k+1}$ are set to their mean
  values to ensure that $\hat{\mat{R}}_{k+1}$ and $\hat{\mat{T}}_{k+1}$ are Toeplitz structured.

@@

	\begin{subequations}
		\begin{equation}
			\beta_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\hat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
        - \frac{2}{C}\left[\text{Tr}(\tilde{\mat{S}}_{k+1})\right]^2
        + \frac{1}{C}\text{Tr}(\tilde{\mat{S}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i\hat{\mat{T}}_k^+\mat{X}_i^\intercal)^2\right]
			}
			\label{eq:stbf-struct/spatial-shrinkage}
		\end{equation}
		\begin{equation}
			\gamma_{k+1} =
			1-
			\frac{
        \frac{N}{N-1}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
				+ \frac{1}{S}\text{Tr}(\tilde{T}_{k+1}^2)
				- \frac{1}{N(N-1)}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\mat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}{
        \frac{N^2-2N}{(N-1)^2}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
        - \frac{2}{S}\left[\text{Tr}(\tilde{\mat{T}}_{k+1})\right]^2
        + \frac{1}{S}\text{Tr}(\tilde{\mat{T}}_{k+1}^2)
				+ \frac{1}{N(N-1)^2}\sum_{i=1}^N
        \left[\text{Tr}(\mat{X}_i^\intercal\hat{\mat{S}}_k^+\mat{X}_i)^2\right]
			}
			\label{eq:stbf-struct/temporal-shrinkage}
		\end{equation}
	\end{subequations}
	The shrinkage parameters $0<\beta_{k+1}<1$ and $0<\gamma_{k+1}<1$ should be
	re-determined after each iteration.
	The Oracle Approximation Shrinkage method can also be used to determine
	$\beta_{k+1}$ and $\gamma_{k+1}$~\cite{Chen2010,Xie2021} but performs worse for spatiotemporal \ac{eeg} data since not all assumptions are met.

@@

	%	identity matrix and assuming identity covariance in \cref{eq:stbf-struct/closed-form} yields weights
